Stable individual differences in strategies within, but not between, visual search tasks

### 2019-05-28 Manuscript/manuscript.tex
tidied up a few bits

Stimuli consisted of arrays of black oriented line segments against a grey background. The target was oriented $45^{\circ}$ clockwise, while the distractor items had a random orientation with a mean of $45^{\circ}$ anti-clockwise. The variance was low ($18^{\circ}$) on one half of the display to create a homogeneous texture, and high ($95^{\circ}$) on the other side to create a heterogeneous texture. When the target is present on the homogeneous half, it can be easily be detected with peripheral vision, but when it is in the heterogeneous half, it is much harder to detect (this was verified in \cite{nowakowska2017}). There were 160 trials in total and homo- and heterogeneous sides of the display were randomly varied from trial to trial. The dominant eye position was recorded using a desktop-mounted EyeLink 1000 eye tracker (SR Research, Canada).

@@

Each search display was composed of red, blue, green and variable-coloured small squares arranged in three concentric rings around fixation. Variable distractors changed colours from trial-to-trial according to a 24 trial cyclical pattern: the distractors would be red for five trials, then across a period of seven trials, they would gradually change colour from red to blue. The variable distractor would then be blue for five trials, and then gradually transition back to red.

A white digit appeared inside each square. Participants were informed that two targets - a red square and a blue square each with a digit between 2 and 5 - were embedded in every search display. The two target digits were always different, to enable us to distinguish the chosen target. The remaining red, blue and variable squares all contained digits between 6-9. Green squares could contain any digit between 2-9. The location of the targets and distractor within the search display were randomized on each trial. Participants were only required to find one target on each trial, and they were free to search for either one.

@@

We successfully replicated the wide range of individual differences in strategy and performance that had previously been observed in each of these three visual search paradigms, with a larger sample size than the original experiments. Surprisingly, however, the between-paradigm correlations give $R^2 \approx 0.04$, and this is a generous interpretation that would fail to pass the usual criteria for null hypothesis significance testing. Knowing how one person will behave in one of these paradigms tells us very little about how they will perform in the others. This can be contrasted with the relatively high test-retest correlations of all three of the tasks individually. Indeed, we find that the visual search tasks under consideration here compare favourably to other cognitive psychology paradigms, such as the Erkison Flanker and Posner Cueing tasks in terms of test-retest reliability \cite{hedge2018}.

There are many reasons why two measurements might be uncorrelated, such as range restriction or measurement noise. The test-retest correlations on each of the individual visual search tasks rule out many of these alternatives, leaving a true absence of shared variance between these tasks as a likely explanation. One might have expected reaction time to be at least modestly correlated from one search task to the next, as a general factor like an individual's speed-accuracy trade-off, or motivation might lead to better or worse overall performance. However, we did observe correlations between measures of strategy and reaction time within each task, suggesting our strategy metrics determine a large proportion of search performance.

Although the tasks in this experiment all have visual search in common, they also have unique aspects that appear to have resonated with particular individuals' strengths, and not others. Our definition of a successful strategy in the SHLS task was fixating the locations that provide new information. In the ACVS task, a successful strategy meant appropriately altering search goals to match changes in the environment. In the MCFT task, success involved minimizing cognitive load by minimizing target switching. Each of these tasks taps into unique aspects of visual search strategies, and performance on one has little bearing on the others.

>>>>

Stimuli consisted of arrays of black oriented line segments against a grey background. The target was oriented $45^{\circ}$ clockwise, while the distractor items had a random orientation with a mean of $45^{\circ}$ anti-clockwise. The variance was low ($18^{\circ}$) on one half of the display to create a homogeneous texture, and high ($95^{\circ}$) on the other side to create a heterogeneous texture. When the target is present on the homogeneous half, it can be easily be detected with peripheral vision, but when it is in the heterogeneous half, it is much harder to detect. This was verified in \cite{nowakowska2017}: for brief presentations viewed from the center, detection performance was close to chance for targets presented on the heterogeneous texture, and close to ceiling for targets presented on the homogeneous texture). There were 160 trials in total and homo- and heterogeneous sides of the display were randomly varied from trial to trial. The dominant eye position was recorded using a desktop-mounted EyeLink 1000 eye tracker (SR Research, Canada).

@@

Each search display was composed of 54 red, blue, green and variable-coloured small squares arranged in three concentric rings around fixation (see \ref{fig:exampleStimuli}). Variable distractors changed colours from trial-to-trial according to a 24 trial cyclical pattern: these distractors would be red for five trials, then across a period of seven trials, they would gradually change colour from red to blue. The variable distractors would then be blue for five trials, and then gradually transition back to red. 
%A missing detail reviewers might want to know here is how many distractors are variable vs fixed colors

A white digit appeared inside each square. Participants were informed that two targets - a red square and a blue square each with a digit between 2 and 5 - were embedded in every search display. The two target digits were always different, to enable us to distinguish the color of the target that had been found on each trial. The remaining red, blue and variable squares all contained digits between 6-9. Green squares could contain any digit between 2-9. The location of the targets and distractor within the search display were randomized on each trial. Participants were only required to find one target on each trial, and they were free to search for either one.

@@

We successfully replicated the wide range of individual differences in strategy and performance that had previously been observed in each of these three visual search paradigms, with a larger sample size than the original experiments. Surprisingly, however, the between-paradigm correlations give $R^2 \approx 0.04$; even a generous interpretation of the correlation between tasks would fail to pass the usual criteria for null hypothesis significance testing. Knowing how one person will behave in one of these paradigms tells us very little about how they will perform in the others. This lack of any consistent relationship between the search tasks occured despite the relatively high test-retest correlations of all three of the tasks individually. Indeed, the test-retest reliability of each of the three measures of visual search strategy we used in this study compare favourably to other cognitive psychology paradigms, such as the Erkison Flanker and Posner Cueing tasks, making them well suited for detecting relationships with other variables \cite{hedge2018}. We also observe strong correlations between measures of strategy and reaction time \textit{within} each task. These correlations demonstrate that our strategy metrics determine a large proportion of search performance, and that our measurements are sufficiently reliable to produce clear correlations where they exist.

There are many reasons why two measurements might be uncorrelated, such as range restriction or measurement noise, but the test-retest correlations and within-task correlations on each of the individual visual search task metrics rule out many of these alternatives, leaving a true absence of shared variance between these tasks as a likely explanation for the lack of correlation. One might have expected reaction time to be at least modestly correlated from one search task to the next, as a general factor like an individual's speed-accuracy trade-off, or motivation might lead to better or worse overall performance, but there was no relationship. Although the tasks in this experiment all have visual search in common, they also have unique aspects that appear to have resonated with particular individuals' strengths, and not others. Our definition of a successful strategy in the SHLS task was fixating the locations that provide new information. In the ACVS task, a successful strategy meant appropriately altering search goals to match changes in the environment. In the MCFT task, success involved minimizing cognitive load by minimizing target switching. Each of these tasks taps into unique aspects of visual search strategies, and performance on one has little bearing on the others.
